### 2020-02-21 doc/icfp20/counterpoint.tex
remove some code

\newcommand{\Motion}{
  \begin{figure}
    \includegraphics[width=6cm]{fig/motion.png}
    \hspace{1cm}
    \includegraphics[width=3cm]{fig/badmotion.png} \\
    %\vspace{-1cm}
    \begin{flushleft}
    \begin{small}
      \hspace{1.95cm} contrary \ \ \ \ parallel \ \ \ \ \ \ \ similar \ \ \ \ \ \ oblique
      \hspace{2.2cm} 5th \hspace{9.5mm} 8th
    \end{small}
    \end{flushleft}
    \caption{Left: Four Kinds of Motion, Right: Invalid Motion}
    \label{fig:motion}
  \end{figure}
}

@@

First species counterpoint is the simplest variant of counterpoint.
In first species, we set one note against each note in the cantus firmus,
which is required to start with a tonic and consist only of whole notes.
Figure~\ref{fig:fs} shows an example of first species counterpoint:

\begin{figure}[h]
  \includegraphics[width=6cm]{fig/fs.png}
  \caption{First Species Counterpoint}
  \label{fig:fs}
\end{figure}

\noindent The lower line is the cantus firmus, which we excerpt from
a famous German song called ``Froschgesang'' (Frog's song).
The upper line is the counterpoint composed by the second author.

@@

In our implementation, we represent each bar as a pitch-interval
pair \texttt{(p ,  i)}, where \texttt{p} is a note in the cantus firmus,
and \texttt{i} is the distance between \texttt{p} and the corresponding
counterpoint note.
We then represent a sequence of bars as a list of pitch-interval pairs,
but with one proviso: we separate the first and last bars from the
middle bars.
Therefore, the Frog's song in Figure~\ref{fig:fs} is represented as a
compound of the following three elements:

@@

The independence of melodic lines is also affected by \emph{motion},
i.e., the way one interval moves to another interval.

\begin{figure}
    \includegraphics[width=6cm]{fig/motion.png}
    \hspace{1cm}
    \includegraphics[width=3cm]{fig/badmotion.png} \\
    \begin{flushleft}
    \begin{small}
      \hspace{1.95cm} contrary \ \ \ \ parallel \ \ \ \ \ \ \ similar \ \ \ \ \ \ oblique
      \hspace{2.2cm} 5th \hspace{9.5mm} 8th
    \end{small}
    \end{flushleft}
    \caption{Four Kinds of Motion (left) and Bad Motion (right)}
    \label{fig:motion}
  \end{figure}

As can be seen from the figure on the left, there are four kinds of
motion: \emph{contrary} (two lines go in different directions),
\emph{parallel} (two lines go in the same direction by the same
distance), \emph{similar} (two lines go in the same direction),
and \emph{oblique} (one line plays the same note).
Among these variants, parallel and similar motion makes it harder
to distinguish between two voices, especially when the second
interval is a perfect one, as in the figure on the right.
To avoid problematic motion, we define the \texttt{checkMotion}
function, which lists all invalid motion approaching a perfect interval.

@@

We next discuss a more complex variant of counterpoint, called the
second species.
In second species, we set \emph{two} half notes against every cantus
firmus note.
Figure~\ref{fig:ss} is an example of such two-against-one counterpoint,
again composed for the Frog's song:

\begin{figure}[h]
  \includegraphics[width=7cm]{fig/ss.png}
  \caption{Second Species Counterpoint}
  \label{fig:ss}
\end{figure}

As we can see from this example, the first and last bars may have
a different structure from middle bars.
More specifically, it is encouraged to begin the counterpoint line with
a half rest and end with a whole note.
Therefore, in our implementation, we reuse the \texttt{PitchInterval} 
type for the first and last bars, and define a new type
\texttt{PitchInterval2} for middle bars:

>>>>

\newcommand{\Motion}{
  \begin{figure}
    \includegraphics[width=6cm]{fig/motion.png}
    \hspace{1cm}
    \includegraphics[width=3cm]{fig/badmotion.png} \\
    \begin{flushleft}
    \begin{small}
      \hspace{1.95cm} contrary \ \ \ \ parallel \ \ \ \ \ \ \ similar \ \ \ \ \ \ oblique
      \hspace{2.2cm} 5th \hspace{9.5mm} 8th
    \end{small}
    \end{flushleft}
    \caption{Left: Four Kinds of Motion, Right: Invalid Motion}
    \label{fig:motion}
  \end{figure}
}

@@

\FS

First species counterpoint is the simplest variant of counterpoint.
In first species, we set one note against each note in the cantus firmus,
which is required to start with a tonic and consist only of whole notes.
Figure~\ref{fig:fs} shows an example of first species counterpoint.
The lower line is the cantus firmus, which we excerpt from a famous
German song called ``Froschgesang'' (Frog's song).
The upper line is the counterpoint composed by the second author.

@@

In our implementation, we represent each bar as a pitch-interval pair
\texttt{(p ,  i)}, where \texttt{p} is the pitch of a cantus firmus note,
and \texttt{i} is the interval between \texttt{p} and the corresponding
counterpoint note.
We then represent a sequence of bars as a list of pitch-interval pairs,
but with one proviso: we separate the first and last bars from the
middle bars.
Therefore, the Frog's song in Figure~\ref{fig:fs} is represented as a
compound of the following three elements:

@@

\Motion

The independence of melodic lines is also affected by \emph{motion},
i.e., the way one interval moves to another interval.
As can be seen from Figure~\ref{fig:motion} (left), there are four
kinds of motion: \emph{contrary} (two lines go in different directions),
\emph{parallel} (two lines go in the same direction by the same
distance), \emph{similar} (two lines go in the same direction),
and \emph{oblique} (one line plays the same note).
Among these variants, parallel and similar motion makes it harder
to distinguish between two voices, especially when the second
interval is a perfect one, as in Figure~\ref{fig:motion} (right).
To avoid problematic motion, we define the \texttt{checkMotion}
function, which lists all invalid motion approaching a perfect interval.

@@

\SS

We next discuss a more complex variant of counterpoint, called the
second species.
In second species, we set \emph{two} half notes against every cantus
firmus note.
Figure~\ref{fig:ss} is an example of such two-against-one counterpoint,
again composed for the Frog's song.
As we can see from this example, the first and last bars may have
a different structure from middle bars.
More specifically, it is encouraged to begin the counterpoint line with
a half rest and end with a whole note.
Therefore, in our implementation, we reuse the \texttt{PitchInterval} 
type for the first and last bars, and define a new type
\texttt{PitchInterval2} for middle bars:

### 2020-02-21 doc/icfp20/music.tex
figures
move figures to fig/
remove some code

\paragraph{Pitches}
In music, a \emph{pitch} is a value denoting how high or low a sound is.
Pitches can be represented either as an absolute value or as a relative
value paired with the octave the sound belongs to.
In our implementation, the two representations of pitches are defined
as \texttt{Pitch} and \texttt{PitchOctave}, respectively, accompagnied
by a proof that converting back and forth between them is an identity.

\begin{alltt}
data Pitch : Set where
  pitch : \(\mathbb{N}\) \(\rightarrow\) Pitch

data Octave : Set where
  octave : \(\mathbb{N}\) \(\rightarrow\) Octave

PitchOctave : Set
PitchOctave = Fin 12 \(\times\) Octave
\end{alltt}

@@

\begin{alltt}
data Duration : Set where
  duration : \(\mathbb{N}\) \(\rightarrow\) Duration
\end{alltt}

\paragraph{Note}
Combining pitches and duration gives us \emph{notes}.
In our implementation, we represent notes with sound as \texttt{tone}s
and those without as \texttt{rest}s.

\begin{alltt}
data Note : Set where
  tone : Duration \(\rightarrow\) Pitch \(\rightarrow\) Note
  rest : Duration         \(\rightarrow\) Note
\end{alltt}

\paragraph{Intervals}
\emph{Intervals} are yet another key concept in music.
An interval denotes the difference in pitch between two notes.
There are 13 kinds of interval within an octave, and these intervals
may be classified from several different perspectives: (i) major or minor;
(ii) consonant or dissonant; and (iii) perfect or imperfect.
For convenience, we let the names of intervals carry the information
about the first and third classifications.
As a convention, we refer to \texttt{min2} and \texttt{maj2}  uniformly
as ``2nd'', and similarly for other intervals.
We also call \texttt{per1} the \emph{unison} in the rest of this section.

\begin{alltt}
data Interval : Set where
interval : \(\mathbb{N}\) \(\rightarrow\) Interval

per1  = interval 0 -- consonant
min2  = interval 1
maj2  = interval 2
min3  = interval 3 -- consonant
maj3  = interval 4 -- consonant
per4  = interval 5 
aug4  = interval 6
per5  = interval 7 -- consonant
min6  = interval 8 -- consonant
maj6  = interval 9 -- consonant
min7  = interval 10
maj7  = interval 11
per8  = interval 12 -- consonant
\end{alltt}

>>>>

\paragraph{Pitches}
In music, a \emph{pitch} is a value denoting how high or low a sound is.
Pitches can be represented either as an absolute value or as a relative
value paired with the octave the sound belongs to.
In our implementation, the two representations of pitches are defined
as the \texttt{Pitch} and \texttt{PitchOctave} datatypes, respectively,
accompanied by a proof that converting back and forth between them
is an identity.

% \begin{alltt}
% data Pitch : Set where
%   pitch : \(\mathbb{N}\) \(\rightarrow\) Pitch

% data Octave : Set where
%   octave : \(\mathbb{N}\) \(\rightarrow\) Octave

% PitchOctave : Set
% PitchOctave = Fin 12 \(\times\) Octave
% \end{alltt}

@@

% \begin{alltt}
% data Duration : Set where
%   duration : \(\mathbb{N}\) \(\rightarrow\) Duration
% \end{alltt}

\paragraph{Notes}
Combining pitches and duration gives us \emph{notes}.
In our implementation, we represent notes as a datatype \texttt{Note}
with two constructors: \texttt{tone} for notes with sound and
\texttt{rest} for those without.

% \begin{alltt}
% data Note : Set where
%   tone : Duration \(\rightarrow\) Pitch \(\rightarrow\) Note
%   rest : Duration         \(\rightarrow\) Note
% \end{alltt}

\paragraph{Intervals}

\emph{Intervals} are yet another key concept in music.
An interval denotes the difference in pitch between two notes.
There are 13 kinds of interval within an octave, and these intervals
may be classified from several different perspectives: (i) major or minor;
(ii) consonant or dissonant; and (iii) perfect or imperfect.
As a convention, we refer to \texttt{min2} and \texttt{maj2}  uniformly
as ``2nd'', and similarly for other intervals.
We also call \texttt{per1} the \emph{unison} in the rest of this section.

\begin{figure}[h]
  \includegraphics[width=12cm]{fig/interval.png} \\
  \begin{flushleft}
    \begin{footnotesize}
      \hspace{1.45cm} \texttt{per1 \hspace{0.5mm} min2  \hspace{2.5mm}
        maj2 \hspace{1.2mm} min3 \hspace{0.5mm} maj3 \ per4
        \hspace{1mm}aug4
        \hspace{0.5mm}  per5\hspace{1.2mm}  min6\hspace{1.6mm} maj6
        \hspace{1.6mm}min7\hspace{1.2mm} maj7\hspace{1.2mm} per8} \\
      consonant? \hspace{1.5mm} yes \hspace{4mm} no \hspace{7mm}  no
      \hspace{4.5mm} yes \hspace{3.8mm}  yes \hspace{3.5mm} no \hspace{3.5mm}
      no \hspace{4mm} yes \hspace{4mm} yes \hspace{3.2mm} yes
      \hspace{3.7mm} no \hspace{4mm} no \hspace{3.8mm} yes
    \end{footnotesize}
  \end{flushleft}
\end{figure}
